---
title: "Decentralization"
bibliography: decentralizationind.bib
author: "Rose-Camille Vincent, Daniel Jemberu and Giulia Minnucci"
date: "Friday, November 07, 2014"
output: html_document
---

Indicator for Decentralization:
Decentralization typically defined in public planning, management and decision making as the transfer of authority and power from higher to lower levels of governing or from national to sub-national levels[@rondinelli1983collins].it has different characterstics for different writers as describe as delegation,devaluation,de-concentration,and privatization of responsibility and authority of tasks to lower level of administration.
Decentralization system involves two different measures: level and degree .decentralizations may occur at system or organization level(Mintzberg(1979).Degree can be defined as the distribution of power over the decision made inside an organization.
Indicators for decentralization
As defined above, decentralization is a complex multiple phenomenon, encompassing a number of political, fiscal and administrative dimension, it is difficult to capture the process and quantify for empirical study. So it leads to a challenge for measuring the outcome of decentralization on health provision system. 
Despite a number of theoretical frameworks, few try to measure the level and the extent of decentralization based on standard space map. The approach will consider both the fiscal and non fiscal aspects of decentralization. Such as  Share of local income/ spending as a proportion of national income/spending  ,major decentralization healthy program (dummy variable pre and post period),citizen participation (high, medium or low),governance rules on healthy offices(high,medium,or low)
We proposed here to use a Revenue authonomy(RA) measure approch  that, used in the [@akai2002sakata] and [@habibi2003etal]iteratures .This approach will consider both own municipalities revenue and total expenditure.Thus,takes the ratio of sub nationally financed administered revenue to the total expenditure that financed and administered by both central and sub national .
Rhe following part will present and analyse Benin's Revenue authonomy(RA) measure for 77 muncipalities from 2005-2010. 

*Loading Data and cleaning
```{r, echo=FALSE}
Decentdat<-read.csv(file.choose(),header = TRUE)
Decentdat
decentralazation <- na.omit(Decentdat)
decentralazation
```

*Checking
```{r}
head(decentralazation)
names(decentralazation)
tail(decentralazation)
sum(is.na(decentralazation))



```{r}
summary(decentralazation)
mean(decentralazation$Decentralazation)
plot(decentralazation$Decentralazation)


```

#Reference



