---
title: "Index.rmd"
author: "Giulia Minnucci,Daniel Jemberu, Rose-Camille Vincent"
date: "20 novembre 2014"
output: html_document
---





#Introduction



For the past decade, governments and policymakers in developing countries have invested in improving basic sanitation and health-related facilities in urban and rural areas. Decentralization and devolution of public administration are two important channels through which these governments, and development agencies, achieve these goals. 
Health decentralization policies have already been implemented in  countries such as Gana, Zambia, Uganda and on a broader scale all over the world as a step in a process of economic and political change. 
This initiative has also been supported by bilateral and multilateral agencies that have been investing in such projects time and resources. 
The process of decentralization involves a series of mechanisms that are focused on the transferral of fiscal, administrative and political authority from the central level, such as the Ministry of Health, to the institutions at the local level. 
Decentralization is said to have many benefits in the country where it is implemented; for example, improving efficiency in allocating resources to a set of services or improving the use of expenditures, since they are decided locally based on specific needs. 

However, focusing our research on the impacts and effects this process has on health provision in Benin, we want to see if it is truly efficient always, or if it can affect negatively some circumstances and overall the benefits perceived by the population.
For example, we would like to see if the fiscal decentralization somehow resulted in an increase or decrease of health provision, which means the complexity of elements that revolve around health care, like the number of facilities, improved health conditions etc...

Thus, our research question is : which are the effects of decentralization on health provision in Benin? 
A better question would be: **What impact does decentralization have on the performance of the health provision system in Benin?**

Moreover, before proceeding with the research, a basic question needs an answer:

**Why Benin?**

In Benin, legal provisions on decentralization has transfered part of health provision competencies to subnational entities such as districts and municipalities. Since the year 2002, municipal governments are increasingly responsible for providing basic services and infrastructures such as water access. To offset the vertical imbalance of subnational governments, the central government have conceded some tax raising powers and has also use intergovernmental transfers through the National Commission on Local Finances. 

The country is currently ranked 165 out of 187 countries in the 2014 Human Development Index, with a GDP/capita of US$794.433 (5% growth in 2013) and an estimated population of 9.351 million in 2012. For the past decade, decentralization has been at the core of the major national policies, which aims at reducing poverty for a human sustainable development. 
We chose Benin because it is a young democracy and because decentralization, although at its very early stage, was accompanied by a huge transformation of the political and administrative organization. 


```{r, echo=FALSE, message=FALSE}
library(ggmap)
Country_Benin<-qmap(location='benin',zoom=6)
print(Country_Benin)
```





#Project


This project is divided into three sections. 
The first one gives a general overview on health-related expenditures and provision in Benin over the years 2005 to 2010. The datasets are obtained from the World Bank (WB) and World Health Organization (WHO). 

The second part looks at the microeconomic effect of decentralization on basic health-related facilities. For this purpose, we compute an indicator for decentralization, and we use health-related indicators provided in Benin's Integrated Modular Households Survey (EMICoV). 
Since we have missing values in the households survey, we have decided to restrict our model to only two years: 2010 and 2011, for which we have a balanced panel. 
The final section presents our basic regression models and some potential issues.

The datasets are obtained from the Ministry of Economy and Finances, the National Institute of Statistics and the National Commission on Local Finances. 
Additional information or variables will be drawn from international or regional databases such as the World Bank Indicators(WDI), the World Health Organization(WHO) or Afro-barometer.



**World Bank Indicators**

The first dataset is obtained from the World Bank (WB) and includes the following indicators:

(i) health expenditure per capita (SH.XPD.PCAP): sum of public and private health expenditures as a ratio of total population. This indicator covers the provision of health services (preventive and curative), family planning activities, nutrition activities, and emergency aid designated for health but does not include provision of water and sanitation. Data are in current U.S. dollars.

(ii) public health expenditure (SH.XPD.PUBL): recurrent and capital spending from government (central and local) budgets, external borrowings and grants (including donations from international agencies and nongovernmental organizations), and social (or compulsory) health insurance funds. 

(iii) improved sanitation facilities(SH.STA.ACSN): access to improved sanitation facilities as the percentage of the population using improved sanitation facilities. This indicator includes flush/pour flush (to piped sewer system, septic tank, pit latrine), ventilated improved pit (VIP) latrine, pit latrine with slab, and composting toilet.

```{r, include=FALSE}
library(WDI)
healthpublicexpend<-WDI(indicator= c('SH.XPD.PUBL','SH.XPD.PCAP','SH.STA.ACSN'))
Benin<-subset(healthpublicexpend,country=='Benin')
Benin$iso2c<-NULL
names(Benin)[names(Benin)=="year"] <- "Year"
names(Benin)[names(Benin)=="SH.XPD.PUBL"] <- "Public.Health.Expenditure"
names(Benin)[names(Benin)=="SH.XPD.PCAP"] <- "Health.Expenditure.Per.Capita"
names(Benin)[names(Benin)=="SH.STA.ACSN"] <- "Improved.Sanitation.Facilities"

```



**World Health Organization Indicators**

The second dataset is taken from the WHO databases, particularly from the Global Health Obervatory Data which provides information on health infrastructures per 100,000 population in Benin including:

(i) Health posts, that  are either community centres or health environments with a very limited number of beds with limited curative and preventive care resources normally assisted by health workers or nurses,

(ii) Health centers,which includes the number of health centres from the public and private sectors, per 100,000 population

(iii) Number of district/rural hospitals from the public and private sectors, per 100,000 population,

(iv) Number of provincial hospitals from the public and private sectors, per 100,000 population,

(v) Number of specialized hospitals delivering mainly tertiary care from the public and private sectors, per 100,000 population. These specialized hospitals could be regional, specialized, research hospitals or Federal/National Institutes.

(vi) Number of specialized hospitals delivering mainly tertiary care from the public and private sectors, per 100,000 population. These specialized hospitals could be: regional, specialized, research hospitals or Federal/National Institutes.

Unlinke the dataset from the World Bank, this datasets only provide Benin-related indicators for the years 2010 and 2013, which we extracted for this project. 

```{r,include=FALSE}
infrastructures <- read.csv("http://apps.who.int/gho/athena/data/data-text.csv?target=GHO/DEVICES01,DEVICES02,DEVICES03,DEVICES04,DEVICES05&profile=text&filter=COUNTRY:*;REGION:AFR;REGION:AMR;REGION:SEAR;REGION:EUR;REGION:EMR;REGION:WPR")
Benininfra<-subset(infrastructures,Country=='Benin')
Benininfra$PUBLISH.STATES<-NULL
Benininfra$WHO.region<-NULL
Benininfra$Low<-NULL
Benininfra$High<-NULL
Benininfra$Comments<-NULL
library(dplyr)
messy<-data.frame(
  Indicator=c("District/rural hospitals", "Health centres","Provincial hospitals","Specialized hospitals","Provincial hospitals","Specialized hospitals"),
  Year=c(2010,2010,2010,2010,2013,2013),
  Country=c("Benin","Benin","Benin","Benin","Benin","Benin"),
  Display.Value=c(0.29,6.36,0.06,0.12,0.05,0.11)
  )
library(tidyr)
tidy<-spread(messy,Indicator,Display.Value)
```


```{r, include=FALSE}
names(Benin)[names(Benin)=="country"] <- "Country"
total<-merge(Benin,tidy, by=c("Country","Year"),all=TRUE)
```





#Analysis



##Macroeconomic Overview

```{r,include=FALSE, message=FALSE}
library(googleVis)
```

```{r,include=FALSE, results='asis'}
compare_trend <-data.frame(
  Year=c("2005","2006","2007","2008","2009","2010","2011"),
  Public.Health.Expenditure=c(49.69487,50.22141,50.75158,50.95043,53.35198,51.16728,52.10150),
  Health.Expenditure.Per.Capita=c( 25.21300,26.47114,28.77537,31.01608,31.80012,30.47805,33.73303),
  Improved.Sanitation.Facilities=c(11.2,11.7,12.2,12.7,13.2,13.7,14.2))
```

```{r,echo=FALSE, results='asis'}
line_plot <- gvisLineChart((compare_trend), 
              options=list(
                        title="Compare trends",
                        titleTextStyle="{color:'black', 
                                           fontName:'Courier', 
                                           fontSize:16}",   
                          curveType="function",
                          width=900,
                          height=500                         
                          )) 
print(line_plot, tag = 'chart')
```


The above graph suggests that all three elements, public health expenditure, health expenditure per capita, and improved sanitation, have considerably increased over the timespan 2005 to 2011. 
As mentionned above, the central government in Benin (alike other developing countries in West-Africa) has used decentralization as a mechanisms to reach the targets of the Millenium Development Goals.  

Since health expenditure is highly decentralized, we could potentially argue that the macroeconomic overview is related to the outcome on a micro level.






##Microeconomic Analysis 



In this section, we explore provision of health-related services and infrastructures on a local level. Particularly, we look at the impact of decentralization on the access to basic health-related services and infrastructure.
The datasets on local finances are obtained from the National Commission on Local Finances of Benin. Health-related expenditures and provision are from Benin's Integrated Modular Surveys on Household Living Conditions (EMICoV) and the Demographic and Health Survey.

```{r, include=FALSE}
library(RCurl)
library(foreign)  
#url version to be edited 
url <- "https://raw.githubusercontent.com/giumin/Third_Assignment/master/decentral.csv"
decentralization.data <- getURL(url)                
decentralization.data <- read.csv(textConnection(decentralization.data))

```




###Description of Variables 

**Independent variable: Decentralization**

Decentralization is typically defined in public planning, management and decision making, as the transfer of authority and power from higher to lower levels of governing, or from national to sub-national levels. 
It has different characteristics for different writers which often describe it as either delegation, devaluation, de-concentration and privatization of responsibility and authority of tasks to lower level of administration. 

```{r, include=FALSE, results='asis'}

library(stargazer)
```

```{r, results='asis',echo=FALSE}
stargazer(decentralization.data[c("decentralization","population","Density", "pubexp.c")], type = 'html', 
 title="Descriptive statistics/selected variables", digits=2)
```
####scatterplot matrix
The above scatterplot matrix will illustrate the correlation matix between the explanatory variables decentlazation,population,density and public expenditure.In addition,it will show that wether the study varaibles has showed normal distribution or not.
as clearly showed in the graph most of the variables follows positive skewed distribution.
```{r, results='asis',echo=FALSE}
library("car")
car::scatterplotMatrix(decentralization.data[c("decentralization","population","Density", "pubexp.c")])
```

**Dependent variables**

 We have  decided to focus on some key variables which data is only available on a macro level. These variables are drawn from Benin's EmiCoV (2010,2011) and are defined as follwos: 
(i) access to water: defined as the percentage of households in the municipality that have access to water 
(ii)access to toilet facilities: percentage of households in the municipality that have access to toilet facilities 
(iii)water provision infrastructure: percentage of households in that municipality that are connected to SONEB (water pipeline)
(iv) hospital: the number of existing hospitals or public healthcare centers in the municpality. 



**Explanatory variables**

In addition to decentralization, we use a set of variables to account for heterogeneity accross the municipalities, such as population, population density, per capita consumption, public expenditure per capita, households health consumption, indicators on monetary poverty, average literacy rate of the head of households, as well as a monetary poverty index, all of them obtained from Benin's EmiCov (2010, 2011). 
We assume that these variables might have some impact either on health health-related expenditures on municipal or households level, and also some impact on decentralization process per se. 

```{r,include=FALSE}
#URl to be edited 

url2 <- "https://raw.githubusercontent.com/giumin/Third_Assignment/master/health.csv"
health.data <- getURL(url2)                
health.data <- read.csv(textConnection(health.data))

decenthealth <-merge(decentralization.data,health.data, by=c("Year", "ID", "Cities"), all=TRUE)
```



```{r,echo=FALSE, results='asis'}

stargazer(health.data[c("pop_wateraccess","pop_toiletacess","pubhosp","connect_soneb")], type = 'html', title="Descriptive statistics of dependent variables", digits=2)
```
#####scatterplot matrix 
The above scatterplot matrix will illustrate the correlation matix between the study variables water acess,water facilities,water pipeline connection and public healthcare center .In addition,it will show that wether the study varaibles has showed normal distribution or not.
as clearly showed in the graph except acess to water ,water facilities ,both  water pipeline connection and public healthcare center follows positive skewed distribution.
```{r,echo=FALSE, results='asis'}
library("car")
car::scatterplotMatrix(health.data[c("pop_wateraccess","pop_toiletacess","pubhosp","connect_soneb")])
```




###Results

  
**Effect of Decentralization on Water Access**


In this model, decentralization reveals to be insignificant,as well as the other explanatory variables. This may be due to mispecification of the model. 
 
```{r,include=FALSE, results='asis'}

# First regression:water acces
R1water <- lm(pop_wateraccess ~ decentralization, data = decenthealth)
confint(R1water)
summary(R1water)

# Second regression: water access
R2water <- lm(pop_wateraccess ~ decentralization + population + Density, data = decenthealth)
confint(R2water)

# Third regression: water access
R3water <- lm(pop_wateraccess ~ decentralization + population + Density + log(pubexp.c) + log(consump.c) + cons_health + educ_adult + ind_pauv_mon, data = decenthealth)
confint(R3water)

```

 
```{r, echo=FALSE, results='asis'}
labels <- c('Decentralization', 'Population', 'Population Density', 'Log Public Expenditure per capita', 'Log Consumption per Capita', 'Average Households Health Consumption Ratio', 'Average Literacy Rate of Head of Households', 'Monetary Poverty Index','(Intercept)')
stargazer::stargazer(R1water, R2water, R3water, covariate.labels = labels,
                     title = 'Effect of decentralization on water access',
                     digits = 2, type = 'html', header = FALSE)
```




**Effect of decentralization on access to toilet facilities** 


Alike the model on water access, decentralization seems to not have a particular effect on the access to toilet facilities, while average literacy rate of head of households results to be  slightly significant. 


```{r, include=FALSE}
# First regression: toilet access
R1toilet <- lm(pop_toiletacess ~ decentralization, data = decenthealth)
summary(R1toilet)
confint(R1toilet)
# Second regression: toilet access
R2toilet <- lm(pop_toiletacess ~ decentralization + population + Density, data = decenthealth)
summary(R2toilet)
confint(R2toilet)
# Third regression: toilet access
R3toilet <- lm(pop_toiletacess ~ decentralization +  population + Density + log(pubexp.c) + log(consump.c) + cons_health + educ_adult + ind_pauv_mon, data = decenthealth)
summary(R3toilet)
confint(R3toilet)
```

```{r,echo=FALSE,results='asis'}
labels <- c('Decentralization', 'Population', 'Population Density', 'Log Public Expenditure per capita', 'Log Consumption per Capita', 'Average Households Health Consumption Ratio', 'Average Literacy Rate of Head of Households', 'Monetary Poverty Index','(Intercept)')

stargazer::stargazer(R1toilet, R2toilet, R3toilet, covariate.labels = labels,
                     title = 'Effect of decentralization on toilet access',
                     digits = 2, type = 'html', header = FALSE)
```



**Effect of decentralization on water pipeline connection**


In this model, our key explanatory variable 'decentralization' seems to have no impact of connection to water delivery infrastrucure. In reverse, population density, public expenditures per capita and average literacy rate of heads of households reveal to be significant; housholds health consumption is slightly significant. 




```{r,include=FALSE}
# First regression: water pipeline
R1soneb <- lm(connect_soneb ~ decentralization, data = decenthealth)
summary(R1soneb)
confint(R1soneb)
# Second regression: water pipeline
R2soneb <- lm(connect_soneb ~ decentralization + population + Density, data = decenthealth)
summary(R2soneb)
confint(R2soneb)
# Third regression: water pipeline
R3soneb <- lm(connect_soneb ~ decentralization + population + Density + log(pubexp.c) + log(consump.c) + cons_health + educ_adult + ind_pauv_mon, data=decenthealth)
summary(R3soneb)
confint(R3soneb)
```

```{r, echo=FALSE,results='asis'}

labels <- c('Decentralization', 'Population', 'Population Density', 'Log Public Expenditure per capita', 'Log Consumption per Capita', 'Average Households Health Consumption Ratio', 'Average Literacy Rate of Head of Households', 'Monetary Poverty Index','(Intercept)')

stargazer::stargazer(R1soneb, R2soneb, R3soneb, covariate.labels = labels,
                     title = 'Effect of decentralization on water delivery infrastructure',
                     digits = 2, type = 'html', header = FALSE)
```



**Effect of decentralization on availability of public health infrastructure**


Like the previous models, decentralization does not seem to be a significant elements in the availability of healthcare centers at the subnational level. With regards to the explanatory variables: population, population density, and average literacy rate of the head of households seem to be strongly determinants of the existence of healthcare centers. Public expenditures per capita and poverty level result to be significant as well.



```{r,include=FALSE}
# First regression: public healthcare center
R1pubhosp<-lm(pubhosp ~ decentralization, data=decenthealth)
summary(R1pubhosp)
confint(R1pubhosp)
# Second regression: public healthcare center
R2pubhosp <- lm(pubhosp ~ decentralization + population + Density, data = decenthealth)
summary(R2pubhosp)
confint(R2pubhosp)
# Third regression: public healthcare center
R3pubhosp <- lm(pubhosp ~ decentralization + population + Density + log(pubexp.c) + log(consump.c) + cons_health + educ_adult + ind_pauv_mon, data=decenthealth)
summary(R3pubhosp)
confint(R3pubhosp)
```

```{r, echo=FALSE,results='asis'}
stargazer::stargazer(R1pubhosp, R2pubhosp, R3pubhosp, covariate.labels = labels,
                     title = 'Effect of decentralization on availability of healthcare centers',
                     digits = 2, type = 'html', header = FALSE)
```



#Conclusions 

The macroeconomic overview suggests that Benin's has experienced an improvement on health-services delivery over the past few years. Given that health-related services is highly decentralized, we have tested the impact of decentralization on some key health-related variables, issued from Benin's Households Survy. The microeconomic analysis, however, did not find any significant imapct of decentralization (proxied by the ratio of own-revenue to total expenditures) on the provision of health-related services and infrastructures. Given that exisiting literatures provide various decentralization indicators, the significance found in the microeconomic analysis might be the result of a misleading indicator. Moreover, because public services in Benin is highly financed by intergovernmental (rather than own-source revenues), it might be more reliable to consider an indicator or a model that takes into account this factor. Therefore, we propose to revise our proxy for decentralization and to further work on our model specification. 





