---
title: "Presentation"
author: "Daniel Jemberu, Giula Minnucci, Rose Camille Vincent"
date: "Thursday, November 27, 2014"
output: ioslides_presentation
---

## Research Question

 **What impact does decentralization have on the performance of the health provision system in Benin?**
 

## Description of Benin

- GDP per capita: USD 804.67 in 2013

- GDP growth: 5.6% in 2013

- Population: 10.32 million habitants 

- Human Development Index 2014: 165 out of 187 countries 

```{r, echo=FALSE, message=FALSE}
library(ggmap)
Country_Benin<-qmap(location='benin',zoom=6)
print(Country_Benin)
```


##Why Benin?


- Decentralization and devolution have been at the center of national policy for the past decade

- Decentralization: huge transformation of the political and administrative organization

- Decentralization as mechanism to fight poverty and reduce inequality

- Legal provisions has transfered part of health provision competencies to subnational entities 

- Subnational governments have tax raising powers and are allocated transfers to meet assigned expenditures



## Theoretical background

Proponents of decentralization share the view that local governments are the major vehicles for specific poverty alleviation policies. 

**Economic Theory**

- decision-making should occur at the lowest level of government in order to reach allocation efficiency reflecting economies of scale and benefit-cost spill outs [@shah1994reform] 

- local government has an informational advantage, which is essential to improve provision of public services [@oates1972fiscal]

- decentralization shall enhance the accountability of policymakers through greater participation of nearby communities in political decisions [@crook1998democracy]  [@tiebout1956pure]

## Empirical Findings 

**Pros**

- [@santos2005participatory] (Bresil): Decentralization has contributed to double the level of access to basic sanitation and enrollment in elementary schools between 1989 and 1996. Revenue collection increased by 48%. 

- [@bardhan2000capture]: (West Bengal): Fiscal autonomy of local governments expands the volume of service delivery 

- [@faguet2004does] (Bolivia): Public investment in education, water and sanitation rose significantly with decentralization and devolution of administrative authorities

- [@bird1999decentralization] (Asian and Latin-America): Positive effect of decentralization on health, primary education and infrastructure.

- [@galasso2005decentralized] (Bangladesh): Decentralization shifted the balance of power in favor of the poor. 

**Cons**

- [@reinikka2004local] (Uganda): Capture of decentralized school grants by local officials in Uganda. 

- [@treisman2000causes]: (developing countries): more levels of government induce higher perceived corruption, less effective provision of public health services 

- [@del2008local] (Indonesia): decentralization leads to yardstick competition between local governments 


## Methodology and Exploratory Data Collection 

*Macroeconomic approach*

Given that health related services is highly decentralized, there might be a strong correlation between the macro-overview and the microeconomic outcome.
- country's health-related public expenditures (WB, WHO) 
- improved sanitation facilities on a macro level


## Macro overview
```{r, include=FALSE}
library(WDI)
healthpublicexpend<-WDI(indicator= c('SH.XPD.PUBL','SH.XPD.PCAP','SH.STA.ACSN'))
Benin<-subset(healthpublicexpend,country=='Benin')
Benin$iso2c<-NULL
names(Benin)[names(Benin)=="year"] <- "Year"
names(Benin)[names(Benin)=="SH.XPD.PUBL"] <- "Public.Health.Expenditure"
names(Benin)[names(Benin)=="SH.XPD.PCAP"] <- "Health.Expenditure.Per.Capita"
names(Benin)[names(Benin)=="SH.STA.ACSN"] <- "Improved.Sanitation.Facilities"

```

```{r,include=FALSE}
infrastructures <- read.csv("http://apps.who.int/gho/athena/data/data-text.csv?target=GHO/DEVICES01,DEVICES02,DEVICES03,DEVICES04,DEVICES05&profile=text&filter=COUNTRY:*;REGION:AFR;REGION:AMR;REGION:SEAR;REGION:EUR;REGION:EMR;REGION:WPR")
Benininfra<-subset(infrastructures,Country=='Benin')
Benininfra$PUBLISH.STATES<-NULL
Benininfra$WHO.region<-NULL
Benininfra$Low<-NULL
Benininfra$High<-NULL
Benininfra$Comments<-NULL
library(dplyr)
messy<-data.frame(
  Indicator=c("District/rural hospitals", "Health centres","Provincial hospitals","Specialized hospitals","Provincial hospitals","Specialized hospitals"),
  Year=c(2010,2010,2010,2010,2013,2013),
  Country=c("Benin","Benin","Benin","Benin","Benin","Benin"),
  Display.Value=c(0.29,6.36,0.06,0.12,0.05,0.11)
  )
library(tidyr)
tidy<-spread(messy,Indicator,Display.Value)
```

```{r, include=FALSE}
names(Benin)[names(Benin)=="country"] <- "Country"
total<-merge(Benin,tidy, by=c("Country","Year"),all=TRUE)
```

```{r, echo=FALSE}
library(ggplot2)

ggplot(data=Benin,aes(x=Year,y=Public.Health.Expenditure)) + geom_line() + geom_point() +theme_bw(base_size=9) + ggtitle("Trend of total public expenditure over the years\n")
```

---

```{r, echo=FALSE}

ggplot(data=Benin, aes(x=Year,y=Health.Expenditure.Per.Capita)) + geom_line() + geom_point() + theme_bw(base_size = 9) + ggtitle("Trend over the years\n")
``` 
 
---

```{r,echo=FALSE}

ggplot(data=Benin, aes(x=Year,y=Improved.Sanitation.Facilities)) +geom_line() + geom_point() + theme_bw(base_size = 9) + ggtitle("Trend over the years\n")
```


## Micro analysis

- Independent variable: Decentralization

- Dependent variables:
  +access to water: defined as the percentage of households in the municipality that have access to water 
  +access to toilet facilities: percentage of households in the municipality that have access to toilet facilities 
  +water provision infrastructure: percentage of households in that municipality that are connected to SONEB (water pipeline)
  +healthcare centers (hospitals): number of existing hospitals or public healthcare centers in the municpality
    
- Explanatory variables: population, population density, per capita consumption, public expenditure per capita, households health consumption, monetary poverty index, average literacy rate of the head of households, gini coefficient, 

## Results 

```{r,include=FALSE}
url2 <- "https://raw.githubusercontent.com/giumin/Third_Assignment/gh-pages/decenthealth.csv"
combineddata.data <- getURL(url2, ssl.verifypeer = FALSE)                
combineddata.data <- read.csv(textConnection(combineddata.data))
```

```{r, echo=FALSE, results='asis'}
stargazer(combineddata.data[c("pop_wateraccess","pop_toiletacess","pubhosp","connect_soneb")], type = 'html', title="Descriptive statistics of dependent variables", digits=2)
```

**Effect of Decentralization on Water Access**

```{r,include=FALSE, results='asis'}
# First regression:water acces
R1water <- lm(pop_wateraccess ~ decentralization2, data = combineddata.data)
confint(R1water)
summary(R1water)

# Second regression: water access
R2water <- lm(pop_wateraccess ~ decentralization2 + log(population) + Density, data = combineddata.data)
confint(R2water)

# Third regression: water access
R3water <- lm(pop_wateraccess ~ decentralization2 + log(population) + Density + log(pubexp.c) + log(consump.c) + cons_health + educ_adult + ind_pauv_mon + gini + for_dentr, data = combineddata.data)
confint(R3water)
```


```{r, echo=FALSE, results='asis'}
labels <- c('Decentralization', 'Population', 'Population Density', 'Log Public Expenditure per capita', 'Log Consumption per Capita', 'Average Households Health Consumption Ratio', 'Average Literacy Rate of Head of Households', 'Monetary Poverty Index','(Intercept)')
stargazer::stargazer(R1water, R2water, R3water, covariate.labels = labels,
                     title = 'Effect of decentralization on water access',
                     digits = 2, type = 'html', header = FALSE)
```




## Limitations