---
title: "Presentation"
author: "Daniel Jemberu, Giula Minnucci, Rose Camille Vincent"
date: "Thursday, November 27, 2014"
output: ioslides_presentation
---

## Research Question

 **What impact does decentralization have on the performance of the health provision system in Benin?**
 

## Description of Benin

- GDP per capita: USD 804.67 in 2013

- GDP growth: 5.6% in 2013

- Population: 10.32 million habitants 

- Human Development Index 2014: 165 out of 187 countries in

```{r, echo=FALSE, message=FALSE}
library(ggmap)
Country_Benin<-qmap(location='benin',zoom=6)
print(Country_Benin)
```


##Why Benin?
 
- huge transformation of the political and administrative organization

- Decentralization and devolution have been at the center of national policy for the past decade

- Decentralization: huge transformation of the political and administrative organization

- Decentralization as mechanism to fight poverty and reduce inequality

- Legal provisions has transfered part of health provision competencies to subnational entities 



Since the year 2002, municipal governments are increasingly responsible for providing basic services and infrastructures such as water access. To offset the vertical imbalance of subnational governments, the central government have conceded some tax raising powers and has also use intergovernmental transfers through the National Commission on Local Finances. 
-



## Theoretical background


- Bullet 1
- Bullet 2
- Bullet 3

## Data and methodology



## Macro overview
```{r, include=FALSE}
library(WDI)
healthpublicexpend<-WDI(indicator= c('SH.XPD.PUBL','SH.XPD.PCAP','SH.STA.ACSN'))
Benin<-subset(healthpublicexpend,country=='Benin')
Benin$iso2c<-NULL
names(Benin)[names(Benin)=="year"] <- "Year"
names(Benin)[names(Benin)=="SH.XPD.PUBL"] <- "Public.Health.Expenditure"
names(Benin)[names(Benin)=="SH.XPD.PCAP"] <- "Health.Expenditure.Per.Capita"
names(Benin)[names(Benin)=="SH.STA.ACSN"] <- "Improved.Sanitation.Facilities"

```

```{r,include=FALSE}
infrastructures <- read.csv("http://apps.who.int/gho/athena/data/data-text.csv?target=GHO/DEVICES01,DEVICES02,DEVICES03,DEVICES04,DEVICES05&profile=text&filter=COUNTRY:*;REGION:AFR;REGION:AMR;REGION:SEAR;REGION:EUR;REGION:EMR;REGION:WPR")
Benininfra<-subset(infrastructures,Country=='Benin')
Benininfra$PUBLISH.STATES<-NULL
Benininfra$WHO.region<-NULL
Benininfra$Low<-NULL
Benininfra$High<-NULL
Benininfra$Comments<-NULL
library(dplyr)
messy<-data.frame(
  Indicator=c("District/rural hospitals", "Health centres","Provincial hospitals","Specialized hospitals","Provincial hospitals","Specialized hospitals"),
  Year=c(2010,2010,2010,2010,2013,2013),
  Country=c("Benin","Benin","Benin","Benin","Benin","Benin"),
  Display.Value=c(0.29,6.36,0.06,0.12,0.05,0.11)
  )
library(tidyr)
tidy<-spread(messy,Indicator,Display.Value)
```

```{r, include=FALSE}
names(Benin)[names(Benin)=="country"] <- "Country"
total<-merge(Benin,tidy, by=c("Country","Year"),all=TRUE)
```

```{r, echo=FALSE}
library(ggplot2)

ggplot(data=Benin,aes(x=Year,y=Public.Health.Expenditure)) + geom_line() + geom_point() +theme_bw(base_size=9) + ggtitle("Trend of total public expenditure over the years\n")
```

## 

```{r, echo=FALSE}

ggplot(data=Benin, aes(x=Year,y=Health.Expenditure.Per.Capita)) + geom_line() + geom_point() + theme_bw(base_size = 9) + ggtitle("Trend over the years\n")
``` 
 
##
```{r,echo=FALSE}

ggplot(data=Benin, aes(x=Year,y=Improved.Sanitation.Facilities)) +geom_line() + geom_point() + theme_bw(base_size = 9) + ggtitle("Trend over the years\n")
```


## Micro analysis

- Independent variable: Decentralization
- Dependent variables:
    access to water
    -access to toilet facilities
    -water provision infrastructure
    -public health infrastructure
    
- Explanatory variables


## Results 

## Limitations